---
title: "Assignment 4"
subtitle: "GEOS 300, 2024"
  
---



There is no template for this assignment, you can use a .ipynb or .Rmd file from one of your previous assignments as a template to get started.


## Instructions

It is strongly recommended that you complete the assignment in Python or R, although you can complete it in excel/numbers if necessary.  They will download the data for you and get you started with a few blocks of code to make plots and tables.  If you are electing to complete the lab in excel, you can find all the relevant data files here [here](https://github.com/GEOS300/AssignmentData/).  **Note** these data are in .csv format.  If you plan to work in excel, you should save a copy of the data in a suitable excel format (e.g., .xlsx) before making any charts or doing any calculations.

Please upload your answers including all calculations, discussions, tables, and graphs in a single document (PDF or HTML file). Note that you can download your .ipynb file under 'File $\gg$ Download as $\gg$ HTLM or PDF via Latex' and submit this as your report. Label the report document with your name and your student number. Upload your answers to Canvas. **Do not attach a spreadsheet.**

**Make Sure** your student number is include on each plot you produce, and that each plot is using the correct units.


## CA DBB Flux Station

In this lab, assignment, you're going to go back to look at more data from the CA-DBB (aka "BB") Eddy Covariance (EC) Station located in Delta, BC.  We're going to look specifically at a single day of EC observations to consider how the energy and mass flow through the ecosystem over a daily cycle.




The data-set includes half-hourly values of the following variables: absolute temperature $\overline{T}$ at 0.4 m and 2 m above ground in $^{\circC}$, mean horizontal wind $\overline{u}$ at 0.4 and 2 m above ground in $m s^-1$, net all wave radiation $R_n$ at 5 m above ground in $W m^{-2}$, air pressure $P_a$ in $kPa$, photon flux density in $\rm{\mu mol}\,\rm{m}^{-2} \rm{s}^{-1}$ and measured covariances $\overline{u^{\prime}w^{\prime}}$, $\overline{w^{\prime}T^{\prime}}$, $\overline{w^{\prime}\rho_v^{\prime}}$, and $\overline{u^{\prime}\rho_{\rm{CO}_2}^{\prime}}$, all measured at 2 m above ground using an eddy covariance system. Here $u^{\prime}$ and $w^{\prime}$ are instantaneous horizontal and vertical winds in $m s^-1$, respectively, $\rho_v$ is absolute humidity in $\rm{g}\,\rm{m}^{-3}$, and $\rho_{\rm{CO}_2}$ is the molar density of the greenhouse gas carbon dioxide (in $\rm{\mu mol}\,\rm{m}^{-3}$).


Last digit of student number | filename
-----------------------------|---------
0 | 20160620.txt
1 | 20160808.txt
2 | 20160811.txt
3 | 20160812.txt
4 | 20160813.txt
5 | 20160817.txt
6 | 20160729.txt
7 | 20160730.txt
8 | 20160820.txt
9 | 20160821.txt

: The file you are responsible for analyzing, named by date in YYYY (Year) mm (month) dd (day) format.  If you are completing the assignment using Python or R, use one of the example from the previous assignments to format your data download command.  Assignment 2 is a good template; you just need to change three things: 1) use the following url https://raw.githubusercontent.com/GEOS300/AssignmentData/main/CA_DBB_DailyData/20150728.txt 2) update the filename and 3) the date column is named "Date/Time (PST)".  *Note* if you are using R, the default behavior of the language is to substitute all non-alphanumeric characters (slashes, underscores, primes, etc.) with a "." character, so your column names will not perfectly match notation used in the file. You can also view/download your data files [here](https://github.com/GEOS300/AssignmentData/tree/main/CA_DBB_DailyData). {#tbl-dates}



```{r}
# Enter your information below
Student.Name = 'June Skeeter'
Student.Number = 123456789
sprintf('GEOS 300 Assignment Submission for %s: %s',Student.Name,Student.Number)
```




```{r}
library("ggplot2")
library("reshape2")
library("latex2exp")
library("patchwork")
library("tidyr")
library("knitr")
library("plyr")
library("lubridate")
library("zoo")
library("readr")
```


## Question 1

Download your data and print descriptive statistics for all columns.  How do maximum and minimum air temperature at 0.4 m compare to those at 2.0 m?



```{r}
# Insert your code or text answer here.  If writing a text answer in .ipynb, change the cell type to Markdown.  If writing a text answer in .Rmd, just delete the tick marks surrounding the block
```


## Question 2

Convert the air temperature $T$ values at 0.4 m and 2.0 m to potential temperature $\theta$.  Recalculate the descriptive statistics for $\theta_{z=0.4}$ and  $\theta_{z=2}$; how does this adjustment impact the values?



```{r}
# Insert your code or text answer here.  If writing a text answer in .ipynb, change the cell type to Markdown.  If writing a text answer in .Rmd, just delete the tick marks surrounding the block
```


## Question 3

From the data-set provided, calculate $H$, $LE$ and $F_{CO2}$ from the covariances of $\overline{w^{\prime}T^{\prime}}$, $\overline{w^{\prime}\rho_v^{\prime}}$, and $\overline{u^{\prime}\rho_{\rm{CO}_2}^{\prime}}$ respectively.  You can assume that the heat capacity of air is 0.0012 MJ m<sup>-3</sup> K<sup>-1</sup>.  You can estimate the density of air $\rho$ using the Ideal Gas Law @eq-Ideal-Gas-Law and you can estimate the latent heat of vaporization following @eq-Latent-Heat-Vaporization:

$$
PV=nRT
$${#eq-Ideal-Gas-Law}

here,$R$ is the ideal gas constant ($8.31446261815324 Pa\ m^3\ mol^{-1} K ^{-1}$), $T$ is temperature (in Kelvin!), $P$ is air pressure, $V$ is volume, $n$ is the amount of the gas in mols; with $n=m/M$, where $m$ is mass and $M$ is the molar mass.  The molar mass of air is $\approx .028964 kg mol^{-1}$

Following Yao & Rogers (1996), the latent heat of vaporization of water ($L_v$ in $J g^{-1}$) can be approximated as a function of temperature ($T$ in $^{\circ}C$) using a third order polynomial function:

$$
L_v = 2500.8-2.36T-0.0016T^2-0.00006T^3
$${#eq-Latent-Heat-Vaporization}



```{r}
# Insert your code or text answer here.  If writing a text answer in .ipynb, change the cell type to Markdown.  If writing a text answer in .Rmd, just delete the tick marks surrounding the block
```



## Question 4

Create a graph with time of day on the $x$-axis and the three measured flux densities of the surface energy balance ($R_n$, $H$, $H$) on the $y$-axis. Briefly discuss their diurnal course.  Do $H$ and $LE$ sum to $R_n$?  If not, why?



```{r}
# Insert your code or text answer here.  If writing a text answer in .ipynb, change the cell type to Markdown.  If writing a text answer in .Rmd, just delete the tick marks surrounding the block
```



## Question 5

For 16:00, Calculate $H$ based on the aerodynamic method using the simultaneously measured profiles of $\overline{u}$ and $\overline{T}$. How does the value of $H$ compare to $H$ in your answer to question 2? Why could it differ?

## Question 6

Using the covariances and the profile measurements, calculate the eddy diffusivities for sensible heat $K_H$ and momentum $K_M$ at 16:00. Is Reynold's analogy fulfilled or not at 16:00?

## Question 7

For 16:00, calculate the Obukhov-length $L$ and the stability parameter $\zeta = z/L$. Does $\zeta$ qualitatively agree with $Rf$? Is $\zeta = z/L$ really indicating `neutral' as assumed in using the aerodynamic method above?

## Question 8

In the process of photosynthesis, energy is extracted from photons in the PAR range. To assimilate one mole of $\rm{CO}_2$, it requires an energy of 469 kJ. The same amount is released back during respiration (metabolism, decomposition of organic matter). We call this energy flux density the *net biochemical energy storage* $\Delta P$. Calculate $\Delta P$ at 13:00 in $\rm{W}\,\rm{m}^{-2}$. At 13:00, what fraction of $R_n$ is used up in the process of photosynthesis? Is $\Delta P$ a relevant term in the surface energy balance of this bog or not?

## Question 9

Determine the average carbon dioxide ($\rm{CO}_2$) flux density between the bog and the atmosphere for this day and convert it to $\rm{g}\,\rm{CO}_2\,\rm{m^{-2}}\,\rm{day}^{-1}$. Is the surface a net sink or net source for atmospheric $\rm{CO}_2$ on this day?  Note: the molar mass of $\rm{CO}_2$ 44.01 $g mol^{-1}$




